Pulsar Classification using Deep Neural Networks

### 2018-04-23 paper/paper.tex
Moved images folder to the appropriate location. SVGs were made but ran into trouble importing them so they were converted to PDF. Added reference to dataset paper and added their classification accuracy compared to mine.

% Copy/pasta of nips_2017 for future modification found here:
% https://nips.cc/Conferences/2017/PaperInformation/StyleFiles

% if you need to pass options to natbib, use, e.g.:
% \PassOptionsToPackage{numbers, compress}{natbib}
% before loading nips_2017
%
% to avoid loading the natbib package, add option nonatbib:
% \usepackage[nonatbib]{nips_2017}

\usepackage{nips_2017}

% to compile a camera-ready version, add the [final] option, e.g.:
% \usepackage[final]{nips_2017}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{pdfpages}       % Because LaTeX doesn't take SVGs...
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}

@@

\section{Introduction}
  We were tasked with picking a dataset and training a deep learning network. 
  First thing to do was to pick a dataset. Being an open-ended assignment there 
  was uncertainty as to what particular dataset to explore. Initial looks 
  involved different image classification datasets such as CIFAR-10, 
  Caltech 101, and NORB. Down the road there is no desire to work with image 
  data classification, work with files and datasets of continuous and 
  discrete data are preferred. After thorough search of the available datasets 
  it was decided pulsar classification best suited this paper.

@@

  For that endevour scikit-learn was used, specifically the 
  $train\_test\_split()$ method. As is standard for classification problems the 
  data was segmented into 80\% training and 20\% testing data. It was then 
  necessary to place the $class$ label into a data structure for later. For 
  both the train and test data, associated labels are passed, along with the 
  data, as tuples. The psuedocode for this section can be found under Algorithm 
  \ref{alg:input_fn}.
  %
  \begin{algorithm}
    \caption{$input\_fn()$}
    \label{alg:input_fn}
    \begin{algorithmic}
      \STATE $data \gets CSV$\\
      \STATE $test, train \gets train\_test\_split(data)$
      \STATE $test\_label \gets test.pop(class)$
      \STATE $train\_label \gets train.pop(class)$
      \RETURN $(train, train\_label), (test,test\_label)$
    \end{algorithmic}
  \end{algorithm}
  %
\subsection{TensorFlow estimator}
  TensorFlow has built in estimators such as:
  %
  \begin{itemize}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNClassifier}{DNNClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineClassifier}{BaselineClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineRegressor}{BaselineRegressor}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNLinearCombinedClassifier}{DNNLinearCombinedClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearClassifier}{LinearClassifier}
    \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearRegressor}{LinearRegressor}
  \end{itemize}
  %
  You also have the ability to build your own estimator, or classifier, and 
  this is the route that was selected for this project. The custom estimator 
  selected utilizes the $model$ specified at the start of the program, a 
  $model\_dir$ for TensorBoard output which will be discussed in section 
  \ref{sec:tensorboard}, and the $params$ which consist of $feature\_columns$, 
  $hidden\_units$, and $n\_classes$.

@@

\subsection{TensorBoard}\label{sec:tensorboard}
  One of the other reasons for utilizing TensorFlow is the TensorBoard project. 
  TensorBoard allows for visualization of various parameters as defined by the 
  user. Figure \ref{img:acc_smooth} and \ref{img:loss_smooth} are the same results as shown in Figure \ref{img:accuracy} and \ref{img:loss} 
  respectively. The smooothing helps in understanding what may be going on, 
  especially in more turbulant graphs. Another benefit is being able to 
  visualize the graph structure. To use TensorBoard on this project ensure it 
  was installed with TensorFlow and type 
  \[\textnormal{tensorboard\ -\--logdir\ model}\]
  and go to the link it provides (by ctrl clicking or copy/paste).

  \begin{figure}[H]
  \caption{Smoothed Accuracy}\label{img:acc_smooth}
  \begin{center}
  \includegraphics[scale=0.65]{../images/acc}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Smoothed Loss}\label{img:loss_smooth}
  \begin{center}
  \includegraphics[scale=0.65]{../images/loss}
  \end{center}
  \end{figure}

\newpage
\section{Results}\label{sec:results}
  RMSProp gradient descent with 0.001 learning rate, performing batch runs on 
  10\% of the training samples every 100 steps produced an accuracy of 97.8\% 
  averaged over 20 tests. These were taken with 80\% train and 20\% test 
  samples. Other hyperparameters were three fully connected layers with 12 
  nodes each using selu activation.

  \begin{figure}[H]
  \caption{Accuracy}\label{img:term_acc}
  \begin{center}
  \includegraphics[scale=0.75]{../images/accuracy_percent_small}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Accuracy}\label{img:accuracy}
  \begin{center}
  \includegraphics[scale=0.75]{../images/Accuracy}
  \end{center}
  \end{figure}
  %
  \begin{figure}[H]
  \caption{Loss}\label{img:loss}
  \begin{center}
  \includegraphics[scale=0.75]{../images/Loss}
  \end{center}
  \end{figure}

\section{Conclusion}

\section{General formatting instructions}
\label{gen_inst}

The text must be confined within a rectangle 5.5~inches (33~picas)
wide and 9~inches (54~picas) long. The left margin is 1.5~inch
(9~picas).  Use 10~point type with a vertical spacing (leading) of
11~points.  Times New Roman is the preferred typeface throughout, and
will be selected for you by default.  Paragraphs are separated by
\nicefrac{1}{2}~line space (5.5 points), with no indentation.

The paper title should be 17~point, initial caps/lower case, bold,
centered between two horizontal rules. The top rule should be 4~points
thick and the bottom rule should be 1~point thick. Allow
\nicefrac{1}{4}~inch space above and below the title to rules. All
pages should start at 1~inch (6~picas) from the top of the page.

For the final version, authors' names are set in boldface, and each
name is centered above the corresponding address. The lead author's
name is to be listed first (left-most), and the co-authors' names (if
different address) are set to follow. If there is only one co-author,
list both author and co-author side by side.

Please pay special attention to the instructions in Section \ref{others}
regarding figures, tables, acknowledgments, and references.

\section{Headings: first level}
\label{headings}

All headings should be lower case (except for first word and proper
nouns), flush left, and bold.

First-level headings should be in 12-point type.

\subsection{Headings: second level}

Second-level headings should be in 10-point type.

\subsubsection{Headings: third level}

Third-level headings should be in 10-point type.

\paragraph{Paragraphs}

There is also a \verb+\paragraph+ command available, which sets the
heading in bold, flush left, and inline with the text, with the
heading followed by 1\,em of space.

\section{Citations, figures, tables, references}
\label{others}

These instructions apply to everyone.

\subsection{Citations within the text}

The \verb+natbib+ package will be loaded for you by default.
Citations may be author/year or numeric, as long as you maintain
internal consistency.  As to the format of the references themselves,
any style is acceptable as long as it is used consistently.

The documentation for \verb+natbib+ may be found at
\begin{center}
  \url{http://mirrors.ctan.org/macros/latex/contrib/natbib/natnotes.pdf}
\end{center}
Of note is the command \verb+\citet+, which produces citations
appropriate for use in inline text.  For example,
\begin{verbatim}
   \citet{hasselmo} investigated\dots
\end{verbatim}
produces
\begin{quote}
  Hasselmo, et al.\ (1995) investigated\dots
\end{quote}

If you wish to load the \verb+natbib+ package with options, you may
add the following before loading the \verb+nips_2017+ package:
\begin{verbatim}
   \PassOptionsToPackage{options}{natbib}
\end{verbatim}

If \verb+natbib+ clashes with another package you load, you can add
the optional argument \verb+nonatbib+ when loading the style file:
\begin{verbatim}
   \usepackage[nonatbib]{nips_2017}
\end{verbatim}

As submission is double blind, refer to your own published work in the
third person. That is, use ``In the previous work of Jones et
al.\ [4],'' not ``In our previous work [4].'' If you cite your other
papers that are not widely available (e.g., a journal paper under
review), use anonymous author names in the citation, e.g., an author
of the form ``A.\ Anonymous.''

\subsection{Footnotes}

Footnotes should be used sparingly.  If you do require a footnote,
indicate footnotes with a number\footnote{Sample of the first
  footnote.} in the text. Place the footnotes at the bottom of the
page on which they appear.  Precede the footnote with a horizontal
rule of 2~inches (12~picas).

Note that footnotes are properly typeset \emph{after} punctuation
marks.\footnote{As in this example.}

\subsection{Figures}

All artwork must be neat, clean, and legible. Lines should be dark
enough for purposes of reproduction. The figure number and caption
always appear after the figure. Place one line space before the figure
caption and one line space after the figure. The figure caption should
be lower case (except for first word and proper nouns); figures are
numbered consecutively.

You may use color figures.  However, it is best for the figure
captions and the paper body to be legible if the paper is printed in
either black/white or in color.
\begin{figure}[h]
  \centering
  \fbox{\rule[-.5cm]{0cm}{4cm} \rule[-.5cm]{4cm}{0cm}}
  \caption{Sample figure caption.}
\end{figure}

\subsection{Tables}

All tables must be centered, neat, clean and legible.  The table
number and title always appear before the table.  See
Table~\ref{sample-table}.

Place one line space before the table title, one line space after the
table title, and one line space after the table. The table title must
be lower case (except for first word and proper nouns); tables are
numbered consecutively.

Note that publication-quality tables \emph{do not contain vertical
  rules.} We strongly suggest the use of the \verb+booktabs+ package,
which allows for typesetting high-quality, professional tables:
\begin{center}
  \url{https://www.ctan.org/pkg/booktabs}
\end{center}
This package was used to typeset Table~\ref{sample-table}.

\begin{table}[t]
  \caption{Sample table title}
  \label{sample-table}
  \centering
  \begin{tabular}{lll}
    \toprule
    \multicolumn{2}{c}{Part}                   \\
    \cmidrule{1-2}
    Name     & Description     & Size ($\mu$m) \\
    \midrule
    Dendrite & Input terminal  & $\sim$100     \\
    Axon     & Output terminal & $\sim$10      \\
    Soma     & Cell body       & up to $10^6$  \\
    \bottomrule
  \end{tabular}
\end{table}

\section{Final instructions}

Do not change any aspects of the formatting parameters in the style
files.  In particular, do not modify the width or length of the
rectangle the text should fit into, and do not change font sizes
(except perhaps in the \textbf{References} section; see below). Please
note that pages should be numbered.

\section{Preparing PDF files}

Please prepare submission files with paper size ``US Letter,'' and
not, for example, ``A4.''

Fonts were the main cause of problems in the past years. Your PDF file
must only contain Type 1 or Embedded TrueType fonts. Here are a few
instructions to achieve this.

\begin{itemize}

\item You should directly generate PDF files using \verb+pdflatex+.

\item You can check which fonts a PDF files uses.  In Acrobat Reader,
  select the menu Files$>$Document Properties$>$Fonts and select Show
  All Fonts. You can also use the program \verb+pdffonts+ which comes
  with \verb+xpdf+ and is available out-of-the-box on most Linux
  machines.

\item The IEEE has recommendations for generating PDF files whose
  fonts are also acceptable for NIPS. Please see
  \url{http://www.emfield.org/icuwb2010/downloads/IEEE-PDF-SpecV32.pdf}

\item \verb+xfig+ "patterned" shapes are implemented with bitmap
  fonts.  Use "solid" shapes instead.

\item The \verb+\bbold+ package almost always uses bitmap fonts.  You
  should use the equivalent AMS Fonts:
\begin{verbatim}
   \usepackage{amsfonts}
\end{verbatim}
followed by, e.g., \verb+\mathbb{R}+, \verb+\mathbb{N}+, or
\verb+\mathbb{C}+ for $\mathbb{R}$, $\mathbb{N}$ or $\mathbb{C}$.  You
can also use the following workaround for reals, natural and complex:
\begin{verbatim}
   \newcommand{\RR}{I\!\!R} %real numbers
   \newcommand{\Nat}{I\!\!N} %natural numbers
   \newcommand{\CC}{I\!\!\!\!C} %complex numbers
\end{verbatim}
Note that \verb+amsfonts+ is automatically loaded by the
\verb+amssymb+ package.

\end{itemize}

If your file contains type 3 fonts or non embedded TrueType fonts, we
will ask you to fix it.

\subsection{Margins in \LaTeX{}}

Most of the margin problems come from figures positioned by hand using
\verb+\special+ or other commands. We suggest using the command
\verb+\includegraphics+ from the \verb+graphicx+ package. Always
specify the figure width as a multiple of the line width as in the
example below:
\begin{verbatim}
   \usepackage[pdftex]{graphicx} ...
   \includegraphics[width=0.8\linewidth]{myfile.pdf}
\end{verbatim}
See Section 4.4 in the graphics bundle documentation
(\url{http://mirrors.ctan.org/macros/latex/required/graphics/grfguide.pdf})

A number of width problems arise when \LaTeX{} cannot properly
hyphenate a line. Please give LaTeX hyphenation hints using the
\verb+\-+ command when necessary.

\subsubsection*{Acknowledgments}

Use unnumbered third level headings for the acknowledgments. All
acknowledgments go at the end of the paper. Do not include
acknowledgments in the anonymized submission, only in the final paper.

\section*{References}

References follow the acknowledgments. Use unnumbered first-level
heading for the references. Any choice of citation style is acceptable
as long as you are consistent. It is permissible to reduce the font
size to \verb+small+ (9 point) when listing the references. {\bf
  Remember that you can go over 8 pages as long as the subsequent ones contain
  \emph{only} cited references.}
\medskip

\small

[1] Alexander, J.A.\ \& Mozer, M.C.\ (1995) Template-based algorithms
for connectionist rule extraction. In G.\ Tesauro, D.S.\ Touretzky and
T.K.\ Leen (eds.), {\it Advances in Neural Information Processing
  Systems 7}, pp.\ 609--616. Cambridge, MA: MIT Press.

[2] Bower, J.M.\ \& Beeman, D.\ (1995) {\it The Book of GENESIS:
  Exploring Realistic Neural Models with the GEneral NEural SImulation
  System.}  New York: TELOS/Springer--Verlag.

[3] Hasselmo, M.E., Schnell, E.\ \& Barkai, E.\ (1995) Dynamics of
learning and recall at excitatory recurrent synapses and cholinergic
modulation in rat hippocampal region CA3. {\it Journal of
  Neuroscience} {\bf 15}(7):5249-5262.

% [4] Bergstra, J.\ \& Bengio, J.\ (2012) Random Search for Hyper-Parameter 
% Optimization. {\it Journal of Machine Learning Research 13} pp.\ 281-305.

>>>>

\usepackage[final]{nips_2017}
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
\usepackage{pdfpages}       % Because LaTeX doesn't take SVGs...
\usepackage{amsmath}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{natbib}

@@

\section{Introduction}
    We were tasked with picking a dataset and training a deep learning network. 
    First thing to do was to pick a dataset. Being an open-ended assignment 
    there was uncertainty as to what particular dataset to explore. Initial 
    looks involved different image classification datasets such as CIFAR-10, 
    Caltech 101, and NORB. Down the road there is no desire to work with image 
    data classification, work with files and datasets of continuous and 
    discrete data are preferred. After thorough search of the available 
    datasets it was decided pulsar classification provided by 
    \citet{orig_pulsar} was best suited this paper.

@@

    For that endevour scikit-learn was used, specifically the 
    $train\_test\_split()$ method. As is standard for classification problems 
    the data was segmented into 80\% training and 20\% testing data. It was 
    then necessary to place the $class$ label into a data structure for later. 
    For both the train and test data, associated labels are passed, along with 
    the data, as tuples. The psuedocode for this section can be found under 
    Algorithm \ref{alg:input_fn}.
    %
    \begin{algorithm}
        \caption{$input\_fn()$}
        \label{alg:input_fn}
        \begin{algorithmic}
            \STATE $data \gets CSV$\\
            \STATE $test, train \gets train\_test\_split(data)$
            \STATE $test\_label \gets test.pop(class)$
            \STATE $train\_label \gets train.pop(class)$
            \RETURN $(train, train\_label), (test,test\_label)$
        \end{algorithmic}
    \end{algorithm}
    %

\subsection{TensorFlow estimator}
    TensorFlow has built in estimators such as:
    %
    \begin{itemize}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNClassifier}{DNNClassifier}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineClassifier}{BaselineClassifier}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/BaselineRegressor}{BaselineRegressor}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/DNNLinearCombinedClassifier}{DNNLinearCombinedClassifier}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearClassifier}{LinearClassifier}
        \item \href{https://www.tensorflow.org/api_docs/python/tf/estimator/LinearRegressor}{LinearRegressor}
    \end{itemize}
    %
    You also have the ability to build your own estimator, or classifier, and 
    this is the route that was selected for this project. The custom estimator 
    selected utilizes the $model$ specified at the start of the program, a 
    $model\_dir$ for TensorBoard output which will be discussed in section 
    \ref{sec:tensorboard}, and the $params$ which consist of 
    $feature\_columns$, $hidden\_units$, and $n\_classes$.

@@

\subsection{TensorBoard}\label{sec:tensorboard}
    One of the other reasons for utilizing TensorFlow is the TensorBoard 
    project. TensorBoard allows for visualization of various parameters as 
    defined by the user. Figure \ref{img:acc_smooth} and \ref{img:loss_smooth} 
    are the same results as shown in Figure \ref{img:accuracy} and 
    \ref{img:loss} respectively. The smooothing helps in understanding what may 
    be going on, especially in more turbulant graphs. Another benefit is being 
    able to visualize the graph structure. To use TensorBoard on this project 
    ensure it was installed with TensorFlow and type 
    %
    \[\textnormal{tensorboard\ -\--logdir\ model}\]
    %
    and go to the link it provides (by ctrl clicking or copy/paste). 
    %
    \begin{figure}[H]  
    \begin{center}
    \includegraphics[scale=0.65]{images/acc}
    \caption{Smoothed accuracy}\label{img:acc_smooth}
    \end{center}
    \end{figure}
    %
    \begin{figure}[H]  
    \begin{center}
    \includegraphics[scale=0.65]{images/loss}
    \caption{Smoothed loss}\label{img:loss_smooth}
    \end{center}
    \end{figure}
    %
\newpage
\section{Results}\label{sec:results}
    RMSProp gradient descent with 0.001 learning rate, performing batch runs on 
    10\% of the training samples every 100 steps produced an accuracy of 97.8\% 
    averaged over 20 tests. These were taken with 80\% train and 20\% test 
    samples. Other hyperparameters were three fully connected layers with 12 
    nodes each using selu activation. Increasing hidden layers led to noisier 
    loss after the large initial drop. The following had no measurable 
    improvement on the classification accuracy:
    \begin{itemize}
        \item Changing from RMSProp to Adagrad
        \item Changing batch size
        \item Modified learning rate
        \item Depth of layers or number of neurons
    \end{itemize}
    %
    \begin{figure}[H]  
    \begin{center}
    \includegraphics[scale=0.735]{images/accuracy_percent_small}
    \caption{Detailed accuracy}\label{img:term_acc}
    \end{center}
    \end{figure}
    %
    \begin{figure}[H]  
    \begin{center}
    \includegraphics[scale=0.75]{images/Accuracy}
    \caption{Accuracy}\label{img:accuracy}
    \end{center}
    \end{figure}
    %
    \begin{figure}[H]
    \begin{center}
    \includegraphics[scale=0.75]{images/Loss}
    \caption{Loss}\label{img:loss}
    \end{center}
    \end{figure}
    %
\section{Conclusion and future work}
    In conclusion, this particular dataset ended up being relatively simple to 
    classify with modern methods. The original paper reported the exact same 
    accuracy of 97.8\% using a Gaussian Hellinger Very Fast Decision Tree. Room 
    for improvement could be a better interface for the CSV files and possible 
    command line arguments. Additionally, the accuracy would occasionally break 
    98\%, but was rare. There is possible room for improvement on the 
    hyperparameters to increase the accuracy consistently beyond 98\%. The loss 
    function dropped off almost immediately, so it could probably deal with 
    fewer hidden layers and/or neurons. Another possible improvement may be 
    adding extra feature layers that are extrapolated from the pulsars (or 
    possibly non-pulsars).
